---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo = TRUE, include = TRUE}
library(readr)
library(dplyr)
library(ggplot2)

df <- read_csv("activity.zip")
```
## What is mean total number of steps taken per day?

###Calculate the total number of steps taken per day
```{r}

total_step_day <- df %>%
dplyr::group_by(date) %>%
dplyr::summarize(total_steps = sum(steps, na.rm = TRUE))

head(total_step_day)
```
###Histogram of number of steps each day
```{r}
hist(total_step_day$total_steps)
```
### Mean and median of total number of steps per dat
```{r}
mean_step_day <- mean(total_step_day$total_steps)
print(paste0("Mean steps: ",mean_step_day))
median_step_day <- median(total_step_day$total_steps)
print(paste0("Median steps: ",median_step_day))
```

## What is the average daily activity pattern?
### interval pattern
```{r}
interval_pattern <- df %>%
    dplyr::group_by(interval) %>%
    dplyr::summarize(mean_steps = mean(steps, na.rm = TRUE))

plot(y =interval_pattern$mean_steps, x = interval_pattern$interval, type = "l")
```

### What 5 minutes interval contain the max number of steps
```{r}
head(interval_pattern %>% 
dplyr::arrange(desc(mean_steps)) %>%
.$interval,1)
```

## Imputing missing values
### Number of missing values
```{r}

print(paste0("Missing values: ",nrow(df %>% dplyr::filter(is.na(steps)))))

```

### Impute missing values
``` {r}
df2 <- df %>% 
    dplyr::mutate(steps = ifelse(is.na(steps), 
        interval_pattern %>% dplyr::filter(interval == interval) %>% .$mean_steps,
        steps))
```

###Calculate the total number of steps taken per day
```{r}

total_step_day2 <- df2 %>%
dplyr::group_by(date) %>%
dplyr::summarize(total_steps = sum(steps, na.rm = TRUE))

head(total_step_day2)
```

###Histogram of number of steps each day
```{r}
hist(total_step_day2$total_steps)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r}
df2 <- df2 %>%
    dplyr::mutate(day = weekdays(date),
    week_day = factor(ifelse(day  %in% c("Saturday","Sunday"), "weekend", "weekday")))
    
interval_pattern2 <- df2 %>%
    dplyr::group_by(interval, week_day) %>%
    dplyr::summarize(mean_steps = mean(steps, na.rm = TRUE))

p <- ggplot(interval_pattern2, aes(x=interval, y = mean_steps)) +
geom_line() +
facet_grid(week_day ~ .)

print(p)
```

